---
title: "DEM, Sampling, and Krigging"
author: "Walker Crane"
date: "2023-07-21"
output: html_document
---

#Home, Selecting property, clipping raster data, creating GeoTiff, plotting Raster with property boundaries
```{r}
# Step 1: Load the required libraries
library(raster)
library(rgdal)
library(ggplot2)
library(scales)
library(sf)
library(sp)
library(terra)
library(rasterVis)

# Step 2: Read the raster data and land parcel polygon shapefile
#raster_path <- "G:/Soil Sci 585/HW2/Soil-585/Homework 2/USGS_one_meter_x69y340_TX_Eastern_B1_2018.tif"
raster_path <- "G:/Soil Sci 585/HW2/Soil-585/Homework 2/USGS_13_n31w097_20191107.tif"
parcel_path <- "G:/Soil Sci 585/HW2/Soil-585/Homework 2/stratmap22-landparcels_48331_milam_202202.shp"

raster_data <- raster(raster_path)
parcel_data <- readOGR(parcel_path)

# Prop_ID values to match
prop_num <- c(71967)

# Use %in% to filter rows
crane_prop <- parcel_data[parcel_data$Prop_ID %in% prop_num, ]

# Step 3: Set the CRS of the shapefile to match the CRS of the raster
proj4string(crane_prop) <- CRS("+init=epsg:4326 +proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
proj4string(crane_prop)
projection(raster_data) <- CRS("+init=epsg:4326 +proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
projection(raster_data)

# Step 4: Convert raster data to a data frame
raster_df <- as.data.frame(raster_data, xy = TRUE)

# Step 5: Convert parcel_data to an sf object
crane_prop_sf <- st_as_sf(crane_prop)

# Step 4: Convert raster data to a data frame
raster_df <- as.data.frame(raster_data, xy = TRUE)
raster_df$Elev_Feet <- raster_df$Layer_1*3.28084

# Step 5: Convert parcel_data to an sf object
crane_prop_sf <- st_as_sf(crane_prop)

# Step 6: Clip the raster using the land parcel polygon
raster_clip <- crop(raster_data, st_buffer(crane_prop_sf, dist = 20))
raster_clip <- raster_clip*3.28084

# Step 7: Save the raster layer to a Geotiff file
writeRaster(raster_clip, filename = "raster_DF_crane_prop.tif", format = "GTiff", overwrite= TRUE)

# Step 8; Load DEM tiff
elevation <- raster("raster_DF_crane_prop.tif")
plot(elevation)
str(elevation)
names(elevation)

# Convert the raster data to a data.frame
elevation_df <- raster::rasterToPoints(elevation)
elevation_df <- as.data.frame(elevation_df)

# Load ggplot2 and ggspatial
library(ggplot2)
library(ggspatial)

# Function to convert decimal degrees to DMS format with cardinal directions
decimal_to_dms_with_directions <- function(decimal_degrees, axis) {
  degrees <- floor(abs(decimal_degrees))
  minutes <- floor((abs(decimal_degrees) - degrees) * 60)
  seconds <- round((abs(decimal_degrees) - degrees - minutes/60) * 3600, 2)
  
  direction <- ifelse(axis == "x", ifelse(decimal_degrees < 0, "W", "E"), ifelse(decimal_degrees < 0, "S", "N"))
  
  return(paste0(degrees, "°", minutes, "'", seconds, "\" ", direction))
}

# Plot the shapefile with ggplot2 and add the raster data as a background layer
ggplot() +
  geom_raster(data = elevation_df, aes(x = x, y = y, fill = raster_DF_crane_prop)) +    # Plot the raster data
  geom_sf(data = crane_prop_sf, color = "black", fill = "transparent") +       # Plot the shapefile
  scale_fill_viridis_c() +                                                              # Use Viridis color scale (you can change this)
  coord_sf(crs = st_crs(crane_prop_sf)) +                                     # Set the coordinate system for the plot
  labs(x = "Longitude", y = "Latitude", fill = "Digital Elevation Model (ft above Sea Level)") +
  scale_x_continuous(labels = function(x) decimal_to_dms_with_directions(x, "x"), expand = c(0, 0)) +  # Use custom labeling for X-axis
  scale_y_continuous(labels = function(y) decimal_to_dms_with_directions(y, "y"), expand = c(0, 0)) +  # Use custom labeling for Y-axis
  theme(axis.title.x = element_text(size = 12, color = "black"),
        axis.text.x = element_text(size = 6.5, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 6.5, color = "black"),
        legend.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
```
#Home, Creating terrain elements, plotting elements
```{r}
#Create DEM
DEM <- raster("raster_DF_crane_prop.tif")

## Calculate the following terrain parameters using "terrain" function: ('Terrain' requires RasterLayer)
## slope, aspect, tpi, tri, roughness, flowdir 
slope <- terrain(DEM, opt=c("slope"), unit="degrees")    
aspect <- terrain(DEM, opt=c("aspect"), unit="degrees")
tpi <- terrain(DEM, opt=c("tpi"))  ## Topographic Position Index
tri <- terrain(DEM, opt=c("tri"))  ## Terrain Ruggedness Index
roughness <- terrain(DEM, opt=c("roughness"))
flowdir <- terrain(DEM, opt=c("flowdir"))


??terrain
## 4. Stack all the terrain parameters including DEM data into a new variable called "terrain"
terrain <- stack(DEM, slope, aspect, tpi, tri, roughness, flowdir)


# Convert the raster data to a data.frame
flowdir_df <- raster::rasterToPoints(flowdir)
flowdir_df <- as.data.frame(flowdir_df)

# Calculate the center coordinates of each cell
flowdir_df$x_center <- flowdir_df$x + 0.5 * res(flowdir)[1]
flowdir_df$y_center <- flowdir_df$y + 0.5 * res(flowdir)[2]

# Create the flow direction plot using ggplot2
ggplot() +
  geom_raster(data = flowdir_df, aes(x = x, y = y, fill = flowdir)) +    # Plot the raster data
  geom_sf(data = expanded_crane_prop_sf, color = "black", fill = "transparent") +       # Plot the shapefile
  geom_segment(data = flowdir_df, aes(x = x_center, y = y_center, xend = x_center + cos(flowdir * pi/180),
                                     yend = y_center + sin(flowdir * pi/180)),
               arrow = arrow(length = unit(0.2, "inches")), color = "blue", size = 0.5) +  # Add arrows for flow direction
  scale_fill_viridis_c() +                                                              # Use Viridis color scale (you can change this)
  coord_sf(crs = st_crs(expanded_crane_prop_sf)) +                                     # Set the coordinate system for the plot
  labs(x = "Longitude", y = "Latitude", fill = "Flow Direction") +
  scale_x_continuous(labels = function(x) decimal_to_dms_with_directions(x, "x"), expand = c(0, 0)) +  # Use custom labeling for X-axis
  scale_y_continuous(labels = function(y) decimal_to_dms_with_directions(y, "y"), expand = c(0, 0)) +  # Use custom labeling for Y-axis
  theme(axis.title.x = element_text(size = 12, color = "black"),
        axis.text.x = element_text(size = 6.5, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 6.5, color = "black"),
        legend.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
```
# Loading raster and parcel data, creating a GeoTiff
```{r}
# Step 1: Load the required libraries
library(raster)
library(rgdal)
library(ggplot2)
library(scales)
library(sf)
library(sp)
library(terra)
library(rasterVis)

# Step 2: Read the raster data and land parcel polygon shapefile
#raster_path <- "G:/Soil Sci 585/HW2/Soil-585/Homework 2/USGS_one_meter_x69y340_TX_Eastern_B1_2018.tif"
raster_path <- "G:/Soil Sci 585/HW2/Soil-585/Homework 2/USGS_13_n31w097_20191107.tif"
parcel_path <- "G:/Soil Sci 585/HW2/Soil-585/Homework 2/stratmap22-landparcels_48331_milam_202202.shp"

raster_data <- raster(raster_path)
parcel_data <- readOGR(parcel_path)

# Prop_ID values to match
prop_num <- c(71967)

# Use %in% to filter rows
crane_prop <- parcel_data[parcel_data$Prop_ID %in% prop_num, ]

# Step 3: Set the CRS of the shapefile to match the CRS of the raster
proj4string(crane_prop) <- CRS("+init=epsg:4326 +proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
proj4string(crane_prop)
projection(raster_data) <- CRS("+init=epsg:4326 +proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0")
projection(raster_data)

# Step 4: Convert raster data to a data frame
raster_df <- as.data.frame(raster_data, xy = TRUE)

# Step 5: Convert parcel_data to an sf object
crane_prop_sf <- st_as_sf(crane_prop)

# Step 4: Convert raster data to a data frame
raster_df <- as.data.frame(raster_data, xy = TRUE)
raster_df$Elev_Feet <- raster_df$Layer_1*3.28084

# Step 5: Convert parcel_data to an sf object
crane_prop_sf <- st_as_sf(crane_prop)

# Step 6: Clip the raster using the land parcel polygon
raster_clip <- crop(raster_data, st_buffer(crane_prop_sf, dist = 20))
raster_clip <- raster_clip*3.28084

# Step 7: Save the raster layer to a Geotiff file
writeRaster(raster_clip, filename = "raster_DF_crane_prop.tif", format = "GTiff", overwrite= TRUE)

# Step 8; Load DEM tiff
elevation <- raster("raster_DF_crane_prop.tif")

# Convert the raster data to a data.frame
elevation_df <- raster::rasterToPoints(elevation)
elevation_df <- as.data.frame(elevation_df)
```

# Creating function to create number of samples per area
```{r}
set_sampling_points <- function(polygon_sf) {
  # Get the total area of the polygon in acres from the "GIS_AREA" column
  total_area_acres <- sum(polygon_sf$GIS_AREA)
  
  # Set the number of sampling points based on the area criteria
  if (total_area_acres <= 1) {
    num_sampling_points <- 10
  } else if (total_area_acres > 1 & total_area_acres <= 5) {
    num_sampling_points <- round(total_area_acres * 2)
  } else if (total_area_acres > 5 & total_area_acres <= 20) {
    num_sampling_points <- round(total_area_acres / 2)
  } else {
    num_sampling_points <- round(total_area_acres)
  }
  
  return(num_sampling_points)
}

# Assuming you have the "crane_prop_sf" data frame with the column "GIS_AREA"
num_sampling_points <- set_sampling_points(crane_prop_sf)

print(num_sampling_points)  # This will give you the number of sampling points based on the area of the polygons and the specified criteria
```

#Creating polygons for pond and and yard
```{r}
# Load ggplot2 and ggspatial
library(ggplot2)
library(ggspatial)

# Function to convert decimal degrees to DMS format with cardinal directions
decimal_to_dms_with_directions <- function(decimal_degrees, axis) {
  degrees <- floor(abs(decimal_degrees))
  minutes <- floor((abs(decimal_degrees) - degrees) * 60)
  seconds <- round((abs(decimal_degrees) - degrees - minutes/60) * 3600, 2)
  
  direction <- ifelse(axis == "x", ifelse(decimal_degrees < 0, "W", "E"), ifelse(decimal_degrees < 0, "S", "N"))
  
  return(paste0(degrees, "°", minutes, "'", seconds, "\" ", direction))
}

# Generate equally spaced sampling points within the property boundaries based on sample number function
sampling_points <- st_sample(crane_prop_sf, type = "regular", size = num_sampling_points)

# Extract longitude (X) and latitude (Y) values from the POINT geometry
sampling_points_df <- as.data.frame(st_coordinates(sampling_points))
names(sampling_points_df) <- c("X", "Y")  # Rename columns for clarity



# Function to convert DMS (Degrees-Minutes-Seconds) to decimal degrees (DD)
dms_to_dd <- function(degrees, minutes, seconds, direction) {
  sign <- ifelse(direction %in% c("S", "W"), -1, 1)
  dd <- sign * (abs(degrees) + minutes / 60 + seconds / 3600)
  return(dd)
}

# Function to convert meters to decimal degrees
meters_to_dd <- function(meters, latitude) {
  # Earth's radius in meters
  earth_radius <- 6378137
  
  # Conversion factor (approximation for small distances)
  meters_per_degree <- (2 * pi * earth_radius) / 360
  
  # Calculate the distance in decimal degrees
  dd <- meters / meters_per_degree
  
  return(dd)
}

# Buffer distance in meters
buffer_distance_meters <- 30

# Example of pond_location in DMS format
lon_deg <- 96
lon_min <- 56
lon_sec <- 36.2
lon_dir <- "W"  # West (use "E" for East)

lat_deg <- 30
lat_min <- 41
lat_sec <- 10
lat_dir <- "N"  # North (use "S" for South)

# Convert DMS to decimal degrees (DD) for longitude and latitude
lon_pond_dd <- dms_to_dd(lon_deg, lon_min, lon_sec, lon_dir)
lat_pond_dd <- dms_to_dd(lat_deg, lat_min, lat_sec, lat_dir)

# Convert buffer distance from meters to decimal degrees
buffer_distance_dd <- meters_to_dd(buffer_distance_meters, lat_pond_dd)

# Create the pond_location as a point in sf format
pond_central_point <- st_point(c(lon_pond_dd, lat_pond_dd))

# Add a buffer of 50 meters around the pond_central_point
pond_buffer <- st_buffer(pond_central_point, dist = buffer_distance_dd, crs = "+init=epsg:4326")

# Extract the x (longitude) and y (latitude) coordinates from the circle
circle_coords <- st_coordinates(pond_buffer)

# Create a dataframe with the circle coordinates
circle_df <- data.frame(lon = circle_coords[, "X"], lat = circle_coords[, "Y"])

pond_boundary_Poly <- Polygon(circle_df)
pond_boundary_Poly_ps <- Polygons(list(pond_boundary_Poly),1)
pond_boundary_Poly_sps <- SpatialPolygons(list(pond_boundary_Poly_ps))
plot(pond_boundary_Poly_sps)
proj4string(pond_boundary_Poly_sps) <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

data = data.frame(f=99.9)
spdf_pond = SpatialPolygonsDataFrame(pond_boundary_Poly_sps,data)
spdf_pond
summary(spdf_pond)

spdf_pond <- st_as_sf(spdf_pond)
spdf_pond <- st_transform(spdf_pond, st_crs(sampling_area))

# # Convert sf polygons to data frames
# pond_boundary_df <- st_coordinates(pond_boundary_sf) %>%
#   as.data.frame() %>%
#   setNames(c("X", "Y"))

# Five points in DMS format
# Replace these values with your DMS coordinates for the five points
lon_deg <- c(96, 96, 96, 96, 96)
lon_min <- c(56, 56, 56, 56,56)
lon_sec <- c(33.32,33.01, 31.66, 29.71, 29.50)
lon_dir <- rep("W", 5)  # West (use "E" for East)

lat_deg <- c(30, 30, 30, 30, 30)
lat_min <- c(41, 41, 41, 41, 41)
lat_sec <- c(8.95, 11.12, 11.59, 10.83, 9.91)
lat_dir <- rep("N", 5)  # North (use "S" for South)

# Convert DMS to decimal degrees (DD) for longitude and latitude for all five points
lon_dd <- dms_to_dd(lon_deg, lon_min, lon_sec, lon_dir)
lat_dd <- dms_to_dd(lat_deg, lat_min, lat_sec, lat_dir)

# Create an sf polygon object from the five points (assumes a closed polygon)
yard_boundary <- as.data.frame(list(rbind(c(lon_dd[1], lat_dd[1]),  # Point 1
                                         c(lon_dd[2], lat_dd[2]),  # Point 2
                                         c(lon_dd[3], lat_dd[3]),  # Point 3
                                         c(lon_dd[4], lat_dd[4]),  # Point 4
                                         c(lon_dd[5], lat_dd[5]),  # Point 5
                                         c(lon_dd[1], lat_dd[1]))))

yard_boundary_Poly <- Polygon(yard_boundary)
yard_boundary_Poly_ps <- Polygons(list(yard_boundary_Poly),1)
yard_boundary_Poly_sps <- SpatialPolygons(list(yard_boundary_Poly_ps))
plot(yard_boundary_Poly_sps)
proj4string(yard_boundary_Poly_sps) <- CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
data = data.frame(f=99.9)
spdf_yard = SpatialPolygonsDataFrame(yard_boundary_Poly_sps,data)
spdf_yard
summary(spdf_yard)

# Set as simple feature
spdf_yard <- st_as_sf(spdf_yard)
spdf_yard <- st_transform(spdf_yard, st_crs(sampling_area))
```

# Defining sampling area, sample size and removing yard and pond points
```{r}
# install.packages("spatstat")
# library(spatstat)
# spatstat.random::rThomas()

# Define the sampling area within the property boundaries (can be the entire property or a specific region)
sampling_area <- crane_prop_sf
sampling_area <- st_transform(sampling_area, st_crs(crane_prop_sf))

# Now both sampling_area and pond_boundary_Poly_sps have the same CRS, and st_difference should work without any errors.
sampling_area_excluded <- st_difference(sampling_area, spdf_yard)
sampling_area_excluded <- st_difference(sampling_area_excluded, spdf_pond)
plot(sampling_area_excluded)

# Generate sampling points within the updated sampling area (excluding pond and yard)
sampling_points_final <- st_sample(sampling_area_excluded, type = "regular", size = num_sampling_points)
??st_sample
# Convert the sampling points to a data.frame with separate columns for longitude (X) and latitude (Y)
sampling_points_df_final <- as.data.frame(st_coordinates(sampling_points_final))
names(sampling_points_df_final) <- c("X", "Y")  # Rename columns for clarity

# Add DMS coordinates to sampling_points_df_final data frame
sampling_points_df_final$Longitude_DMS <- decimal_to_dms_with_directions(sampling_points_df_final$X, "x")
sampling_points_df_final$Latitude_DMS <- decimal_to_dms_with_directions(sampling_points_df_final$Y, "y")


# Add a new column "Sample_Number" to the sampling_points_df_final data frame with the sampling numbers
sampling_points_df_final$Sample_Number <- 1:nrow(sampling_points_df_final)

#Create .csv file with sample points, coordinates, and number
NoCentsRanchSamplingPoints <- write.csv(sampling_points_df_final, "NoCentsSamplingPoints.csv")

```

#Plot parcel boundary with pond and yard polygons with sampling points
```{r}
# Calculate the centroid of the polygon
yard_centroid <- st_centroid(yard_boundary_sf)

yard_centroid_df <- st_coordinates(yard_centroid) %>%
  as.data.frame() %>%
  setNames(c("X", "Y"))

# Calculate the centroid of the polygon
pond_centroid <- st_centroid(pond_boundary_sf)

pond_centroid_df <- st_coordinates(pond_centroid) %>%
  as.data.frame() %>%
  setNames(c("X", "Y"))

# Plot the shapefile with ggplot2 and add the raster data as a background layer
NoCentsRanch <- ggplot() +
  geom_raster(data = elevation_df, aes(x = x, y = y, fill = raster_DF_crane_prop)) +    # Plot the raster data
  geom_sf(data = crane_prop_sf, color = "black", fill = "transparent") +       # Plot the shapefile
  geom_polygon(data = yard_boundary_df, aes(x = X, y = Y), fill = "orange", alpha = 0.5) +
  geom_text(data = yard_centroid_df, aes(x = X, y = Y, label = "YARD"), color = "black") +
  geom_polygon(data = pond_boundary_df, aes(x = X, y = Y), fill = "lightblue", alpha = 0.5) +
  geom_text(data = pond_centroid_df, aes(x = X, y = Y, label = "POND"), color = "black") +
  geom_point(data = sampling_points_df_final, aes(x = X, y = Y), color = "red", size = 2) +  # Plot the final sampling points
  geom_text(data = sampling_points_df_final, aes(x = X, y = Y, label = Sample_Number), 
            color = "black", vjust = 0, hjust = -0.5) +                # Adding labels to the sampling points
  scale_fill_viridis_c() +                                                              # Use Viridis color scale (you can change this)
  coord_sf(crs = st_crs(crane_prop_sf)) +                                     # Set the coordinate system for the plot
  labs(x = "Longitude", y = "Latitude", fill = "Elevation (ft)", title = "No Cents Ranch DEM w/ Sampling Locations") +
  scale_x_continuous(labels = function(x) decimal_to_dms_with_directions(x, "x"), expand = c(0, 0)) +  # Use custom labeling for X-axis
  scale_y_continuous(labels = function(y) decimal_to_dms_with_directions(y, "y"), expand = c(0, 0)) +  # Use custom labeling for Y-axis
  theme(axis.title.x = element_text(size = 12, color = "black"),
        axis.text.x = element_text(size = 8, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.text.y = element_text(size = 8, color = "black"),
        legend.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position = "right",
        legend.justification = "center", 
        plot.title = element_text(hjust = 0.5))
NoCentsRanch

ggsave("NoCentsRanch.jpeg", NoCentsRanch , dpi = 300)
```

